---
title: "R Cheat Sheets & Notes"
---

## Cheat Sheets

* [R Colors](http://www.stat.columbia.edu/~tzheng/files/Rcolor.pdf)

* [RStudio Cheat Sheets](https://www.rstudio.com/resources/cheatsheets/)

* [R Base Commands Cheat Sheet](https://iqss.github.io/dss-workshops/R/Rintro/base-r-cheat-sheet.pdf)

* [Keyboard Shortcuts](https://support.rstudio.com/hc/en-us/articles/200711853-Keyboard-Shortcuts)

## Notes

Add your own notes here...

```{r}
# Add R commands you want to remember here...

```

```{r}
# Or create other R chunks to put them in like this...

```

```{r, eval=FALSE}
# Using ```{r, eval=FALSE} turns off the chunk, but still shows it.
# Useful when you want to remember code, but not run it in this file.
```




```{r, eval=FALSE}
# Permutation Testing for independent samples...
# Create the data:
set.seed(1140411)
sample1 <- rnorm(30, 69, 2.5)
sample2 <- rnorm(30, 69, 2.5)
theData <- data.frame(values = c(sample1,sample2), group = rep(c(1,2), each=30))
View(theData)
boxplot(values ~ group, data = theData)

# Run the permutation test:

myTest <-  ...
observedTestStat <- myTest$...

observedTestStat

N <- 2000      
permutedTestStats <-  rep(NA, N)
for  (i in 1:N ) {
   permutedTest <- ...
   permutedTestStats[i]  <-  permutedTest$...
}
hist(permutedTestStats)
abline(v=observedTestStat)
sum(permutedTestStats >= observedTestStat)/N
sum(permutedTestStats <= observedTestStat)/N



# Permutation testing for paired data...
# Create the data:
set.seed(121)
sample1 <- rnorm(30, 185, 8)
sample2 <- sample1 - rnorm(30, 0, 3.5)
theData <- data.frame(values = c(sample1,sample2), group = rep(c(1,2), each=30), id = rep(c(1:30),times=2))
View(theData)
with(theData, hist(values[group==1] - values[group==2]))

# Perform the permutation test:

myTest <-  ...
observedTestStat <- myTest$...

N <- 2000      
permutedTestStats <-  rep(NA, N)
for  (i in 1:N ) {
   permutedData <- ...
   permutedTest <- ...
   permutedTestStats[i]  <-  permutedTest$...
}
hist(permutedTestStats)
abline(v=observedTestStat)
sum(permutedTestStats >= observedTestStat)/N
sum(permutedTestStats <= observedTestStat)/N
```




