---
title: "Food at College"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
---

```{r, warning=FALSE, message=FALSE}
# From your file menu (at the top of R-Studio) select:
# "Session -> Set working directory -> To source file location"
# Then play this chunk to get the data into R.
library(mosaic)
library(car)
library(DT)
library(pander)
library(readr)
library(tidyverse)
food <- read_csv("../../Data/food.csv") #food.csv is in the Data folder...
```


----

<!-- Instructions:

 Use the Food dataset and a Kruskal-Wallis Test(s) to answer an interesting question(s) that you come up with.
 
 About the dataset:
 
  * GPA - numerical, self-reported current college GPA

  
  * weight - numerical, self-reported weight in lbs


  * gender - "Female", "Male"


  * breakfast - "Cereal", "Donut" 
      The participants were shown a picture of "Cereal" and a "Donut" and asked which one of these pictures they associate with the word "breakfast."


  * calories_chicken - guessing calories in chicken piadina from the options of:
     1 - 265 
     2 - 430 
     3 - 610 
     4 - 720
    (the variable shows the actual number of calories participants selected)


   * calories_day - Importance of consuming calories per day
     1 - i dont know how many calories i should consume 
     2 - it is not at all important 
     3 - it is moderately important 
     4 - it is very important


   * comfort_food - List 3-5 comfort foods that come to mind. 
     Open ended 


   * comfort_food_reasons - What are some of the reasons that make you eat comfort food?      (i.e., anger, sadness, happiness, boredom, etc) - list up to three
     Open ended


   * comfort_food_reasons_coded
     1 - stress
     2 - boredom
     3 - depression/sadness
     4 - hunger
     5 - laziness
     6 - cold weather
     7 - happiness 
     8 - watching tv
     9 - none 


    * cook - how often do you cook?
     1 - Every day 
     2 - A couple of times a week 
     3 - Whenever I can, but that is not very often  
     4 - I only help a little during holidays 
     5 - Never, I really do not know my way around a kitchen


    * cuisine - what type of cuisine did you eat growing up?
     1 - American
     2 - Mexican.Spanish
     3 - Korean/Asian
     4 - Indian
     5 - American inspired international dishes
     6 - other


    * diet_current - describe your current diet
     open ended

	
    * diet_current_coded
     1 - healthy/balanced/moderated/
     2 - unhealthy/cheap/too much/random/
     3 - the same thing over and over
     4 - unclear


    * drink - which picture do you associate with the word “drink”?
     1 - orange juice
     2 - soda 


    * eating_changes  - Describe your eating changes since the moment you got into college?
     Open ended 


    * eating_changes_coded
     1 - worse
     2 - better
     3 - the same
     4 - unclear


    * eating_changes_coded1
     1 - eat faster
     2 - bigger quantity
     3 - worse quality 
     4 - same food
     5 - healthier
     6 - unclear
     7 - drink coffee 
     8 - less food
     9 - more sweets
     10 - timing 
     11 - more carbs or snacking
     12 - drink more water
     13 - more variety


    * eating_out - frequency of eating out in a typical week 
     1 - Never 
     2 - 1-2 times 
     3 - 2-3 times 
     4 - 3-5 times 
     5 - every day


    * employment - do you work? 
     1 - yes full time 
     2 - yes part time 
     3 - no
     4  - other


    * ethnic_food - How likely to eat ethnic food 
     1 - very unlikely 
     2 - unlikely 
     3 - neutral 
     4 - likely 
     5 - very likely 


    * exercise - how often do you exercise in a regular week?
     1 - Everyday 
     2 - Twice or three times per week 
     3 - Once a week
     4 - Sometimes 
     5 - Never


    * father_education - 
     1 - less than high school 
     2 - high school degree 
     3 - some college degree 
     4 - college degree 
     5 - graduate degree 


    * father_profession - what is your father profession?
     Open ended


    * fav_cuisine - What is your favorite cuisine?
     Open ended


    * fav_cuisine_coded
     0-none
     1 - Italian/French/greek
     2 - Spanish/mexican
     3 - Arabic/Turkish
     4 - asian/chineses/thai/nepal
     5 - American
     6 - African 
     7 - Jamaican
     8 - indian


    * fav_food - was your favorite food cooked at home or store bought? 
     1 - cooked at home 
     2 - store bought 
     3 - both bought at store and cooked at home


    * food_childhood - what was your favorite childhood food?
     Open ended



    * fries - which of these pictures you associate with word fries? 
     1 - Mcdonald’s fries
     2 - home fries

 
    * fruit_day - How likely to eat fruit in a regular day 
     1 - very unlikely 
     2 - unlikely 
     3 - neutral 
     4 - likely 
     5 - very likely 


    * grade_level – 
     1 - freshman 
     2 -Sophomore 
     3 - Junior 
     4 - Senior
     
      
    * greek_food - How likely to eat greek food when available?
     1 - very unlikely 
     2 - unlikely 
     3 - neutral 
     4 - likely 
     5 - very likely 
     
      
    * healthy_feel – how likely are you to agree with the following statement: “I feel very healthy!” ?
      1 to 10 where 1 is strongly agree and 10 is strongly disagree - scale
      
      
    * healthy_meal – what is a healthy meal? Describe in 2-3 sentences.
      Open ended
      
      
    * ideal_diet – describe your ideal diet in 2-3 sentences
      Open ended
      
      
    * Ideal_diet_coded
      1 – portion control
      2 – adding veggies/eating healthier food/adding fruit	
      3 – balance
      4 – less sugar
      5 – home cooked/organic
      6 – current diet
      7 – more protein
      8 – unclear
      
      
     * income
      1 - less than $15,000 
      2 - $15,001 to $30,000 
      3 - $30,001 to $50,000 
      4 - $50,001 to $70,000 
      5 - $70,001 to $100,000 
      6 - higher than $100,000
      
     * indian_food – how likely are you to eat indian food when available
      1 - very unlikely 
      2 - unlikely 
      3 - neutral 
      4 - likely 
      5 - very likely 
      
      
     * Italian_food – how likely are you to eat Italian food when available?
      1 - very unlikely 
      2 - unlikely 
      3 - neutral 
      4 - likely 
      5 - very likely 
      
      
     * life_rewarding – how likely are you to agree with the following statement: “I feel life is very rewarding!” ?
      1 to 10 where 1 is strongly agree and 10 is strongly disagree - scale
      
      
     * marital_status
      1 -Single 
      2 - In a relationship 
      3 - Cohabiting 
      4 - Married 
      5 - Divorced 
      6 - Widowed
      
      
     * meals_dinner_friend – What would you serve to a friend for dinner?
      Open ended
      
      
     * mothers_education 
      1 - less than high school 
      2 - high school degree 
      3 - some college degree 
      4 - college degree 
      5 - graduate degree
      
      
     * mothers_profession – what is your mother’s profession? 
      
      
     * nutritional_check - checking nutritional values frequency 
      1 - never 
      2 - on certain products only 
      3 - very rarely 
      4 - on most products 
      5 - on everything
      
      
     * on_off_campus – living situation
      1 - On campus 
      2 - Rent out of campus 
      3 - Live with my parents and commute 
      4 - Own my own house
      
      
     * parents_cook - Approximately how many days a week did your parents cook? 
      1 - Almost everyday 
      2 - 2-3 times a week 
      3 - 1-2 times a week 
      4 - on holidays only 
      5 - never
      
      
     * pay_meal_out - How much would you pay for meal out? 
      1 - up to $5.00 
      2 - $5.01 to $10.00 
      3 - $10.01 to $20.00 
      4 - $20.01 to $30.00 
      5 - $30.01 to $40.00 
      6 - more than $40.01 
      
      
     * Persian_food - How likely to eat persian food when available?
      1 - very unlikely 
      2 - unlikely 
      3 - neutral 
      4 - likely 
      5 - very likely 
      
      
     * self_perception_weight - self perception of weight 
      6 - i dont think myself in these terms 
      5 - overweight 
      4 - slightly overweight 
      3 - just right 
      2 - very fit 
      1 - slim 
      
      
     * Which of the two pictures you associate with the word soup?
      1 – veggie soup
      2 – creamy soup
      
      
     * sports - sports – do you do any sporting activity?
      1 - Yes 
      2 - No 
      99 – no answer
      
      
     * thai_food - How likely to eat thai food when available?
      1 - very unlikely 
      2 - unlikely 
      3 - neutral 
      4 - likely 
      5 - very likely 
      
      
     * tortilla_calories - guessing calories in a burrito sandwhich from Chipolte?
      1 - 580 
      2 - 725 
      3 - 940 
      4 - 1165
      
     * turkey_calories - Can you guess how many calories are in the foods shown below?
     (Panera Bread Roasted Turkey and Avocado BLT)
      1 - 345 
      2 - 500 
      3 - 690 
      4 - 850
      
      
     * type_sports – what type of sports are you involved?
      Open-ended
      
     * veggies_day - How likely to eat veggies in a day? 
      1 - very unlikely 
      2 - unlikely 
      3 - neutral 
      4- likely 
      5 - very likely
      
     * vitamins – do you take any supplements or vitamins?
      1 – yes
      2 – no
      
     * waffle_calories - guessing calories in waffle potato sandwhich 
      1 - 575 
      2 - 760 
      3 - 900 
      4 - 1315


Note this dataset and description come from: © 2020 Kaggle Inc for further details visit:

[Food choices on Kaggle](https://www.kaggle.com/borapajo/food-choices?select=food_coded.csv)


If you want to give your critiquers some ideas about what you have questions on and would like help with, place those details here.-->




<!--End your comments to your critiquers before this line.-->

----

<!--Begin writing your Analysis after this point.-->

## Background

Disney's Tangled released in 2010, I was only 10 or 11 years old. The film featured the song "Mother knows best", which, characteristic of many of Disney's songs, was quite catchy to sing a long to. My mother always used the phrase (and still does) "Mother knows best" (and she really does) long before we watched the film, thus after it's release, the song quickly became a favorite of hers.

<br/> 

Given this unique data set (food) we can perhaps gain some insight into the one of the affects mothers have on their children. Our variable of interest for this study is students GPA, and they will be separated into groups based on their mothers level of education. The levels of education are separated as follows: <br/>
1 - less than high school <br/>
2 - high school degree <br/>
3 - some college degree <br/>
4 - college degree <br/>
5 - graduate degree <br/>

The study is asking whether students whose mothers sought higher levels of education have higher GPA's than students whose mothers did not (or were not able to) seek higher levels of education. 

We filtered the food data so that all missing values in the GPA and mother_education columns were removed. We also removed any GPA variables that were "non-numeric", in other words "Private" or "Unknown" were the keywords. This way only numeric GPA's remained so that we could convert the data into a numeric data-type for this test.


# {.tabset .tabset-fade .tabset-pills}

## Hide Data

## Show Data

```{r}
# Filter out NA's and non-numeric values in the GPA and mother_education columns...
food.mothers.GPA <- food %>% 
  filter(!is.na(mother_education)) %>% 
  filter(!is.na(GPA) & GPA != "Personal" & GPA != "Unknown") %>% 
  mutate(GPA = as.numeric(GPA))

# Display the data
datatable(food.mothers.GPA[,c("GPA", "mother_education")],
          options = list(lengthMenu = c(5, 10, 15)))
```

# 


<br/>


## Analysis

The following boxplots suggest that mothers without a high school diploma may be more likely to have children with lower GPA's. However, the third and fourth quartiles stretch very high, in fact, they stretch as high as the GPA's across all other groups. This is an interesting observation. Perhaps within our data a mother's level of education can have an impact on their children's GPA at level 1 (Less than high school). A dot plot overlays the first boxplot showing the sample size of that group (5). It is by far the smallest sample size among the groups, this could also be reason for the appearance of our side-by-side boxplots. However, we wish to formally test these assumptions. We will perform a Kruskal-Wallis test to determine if at least one of the distributions is different than the rest. 


```{r}
# Change the spacing of the x and y labels / titles
# mgp = c(xlab/ylab, x-/y-axis labels, x-/y-axis facing in = 0/out = 1)
par(mgp=c(3,1.6,0))

# Display side-by-side boxplots of each group...
boxplot(GPA ~ mother_education, data = food.mothers.GPA, 
        names    = c("Less Than \n High school", "High \n School", "Some \n College", "College \n Degree", "Graduate \n Degree"),
        col      = c("darkslategray1", "darkslategray1", "darkslategray1", "darkslategray1", "darkslategray1"),
        xlab     = "Levels of Mother's Education",
        ylab     = "Student's GPA's",
        main     = "Mothers With a HS Diploma Appear to Have \nSmarter Kids But Likely Only Due to Small Sample Size",
        cex.names= 0.8)

# Subset the smallest group's data for the stripchart
subset_data <- subset(food.mothers.GPA, mother_education == 1)

# Create a vector of 1's with length equal to the number of data points in subset_data
x_coords <- rep(1, length(subset_data$GPA))

# Overlay the first boxplot with the stripchart using the specified x-coordinates
stripchart(subset_data$GPA ~ x_coords, vertical = TRUE, pch = 16, col = "black", add = TRUE)
```

<br/>

The hypotheses for this test, written formally are as follows:

$$
  H_0: \text{All samples are from the same distribution}
$$

$$
  H_a: \text{At least one samples distribution is stochastically different}
$$

These could also be interpreted as, "The GPA's of students is on average the same across all levels of their mother's education", as the null hypothesis, and "Students GPA's are different for at least one level of their mother's education", as the alternative hypothesis. 

<br/>

The significance level for this test will be: 

$$
  \alpha = 0.05
$$



```{r}
# Perform the test and display the results...
pander(kruskal.test(GPA ~ mother_education, data = food.mothers.GPA),
       caption = "Kruskal-Wallis Test of Stochastic Dominance")
```

The P-value of this test (0.4336) is very large. Thus, we fail to reject the null hypothesis (p-val. = 0.4336 > 0.05) because there is insufficient evidence to conclude that the level of a mother's education has an effect on students GPA's. Or, in other words, Students GPA's do not differ on average, across the different levels of their mother's education. 

<br/>

```{r}
# Display the numerical summaries of our data...
favstats(GPA ~ mother_education, data = food.mothers.GPA) %>% 
  pander(caption = "Students GPA Summaries by Mother's Education Level")
```

It's interesting to note the difference in Group 1's First Quartile (2.4) and Median (2.71) GPA's compared to all the other groups. The Third Quartile and the Max GPA's are fairly close or the same across all levels. 

<br/>


## Interpretation

The results of our test suggest that a mothers level education is not sufficient information to determine the GPA's of students (their children). However, the one group that appears to be different (Mothers without a HS diploma) could be due to the small sample size (only n = 5) or even due to chance. Because of this we are unable to determine whether all groups belong to the same distribution or if one really differs. With better data we could perform the test again and see if we find any significant results that would then warrant further action.  


<br/><br/><br/>

