---
title: "Recalling Words"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
---

<script type="text/javascript">
 function showhide(id) {
    var e = document.getElementById(id);
    e.style.display = (e.style.display == 'block') ? 'none' : 'block';
 }
</script>

```{r, include=FALSE}
library(mosaic)
library(car)
library(pander)
library(DT) #You may need to run: install.packages("DT") 
```

```{r, eval=FALSE}
# Play the chunk above and this one to get the data into your Console
View(Friendly)
?Friendly
```


<br />

### Background

Many teachers and other educators are interested in understanding how to best deliver new content to students. In general, they have two choices of how to do this.

1. The Meshed Approach
    * Deliver new content while simultaneously reviewing previously understood content.

2. The Before Approach
    * Deliver new content after fully reviewing previously understood content.

A study was performed to determine whether the *Meshed* or *Before* approaches to delivering content had any positive benefits on memory recall. 

<div style="padding-left:15px;">

##### <a href="javascript:showhide('uniquename')">The Experiment <span style="font-size:8pt;">(click to view)</span></a>


<div id="uniquename" style="display:none;">

Individuals were seated at a computer and shown a list of words. Words appeared on the screen one at a time, for two seconds each, until all words had been shown (40 total). After all words were shown, they were required to perform a few two-digit mathematical additions (like 15 + 25) for 15 seconds to avoid immediate memory recall of the words. They were then asked to write down as many of the 40 words as they could remember. They were given a maximum of 5.3 minutes to recall words.

The process of showing words and recalling words was repeated four times with the same list of words each time (four chances to get it right). The presentation of the first trial was the same for all treatment conditions. However, trials 2, 3, and 4 were slightly different for each treatment condition.

<div style="padding-left:15px;">

The `SFR` group (the control group) stands for Standard Free Recall. In all four trials the same list of 40 words was presented, in a random order each time.

The `Before` group also used the same 40 words during each trial. However, any words that were correctly recalled in a previous trial were presented first, or *before* the words that were not recalled in the last trial. After all the correct words were presented in random order, the non-recalled words were presented in a random order.

The `Meshed` group also used the same 40 words during each trial. However, words that were correctly recalled in a previous trial were alternated with a missed word during the next presentation order. 

</div>

The data records the number of correctly recalled words (out of the 40 possible) from the fourth trial. Results were obtained for 30 students, 10 in each of the three treatment groups: `SFR`, `Before`, and `Meshed`. 

</div>

##### <a href="javascript:showhide('uniquename2')">The Data <span style="font-size:8pt;">(click to view)</span></a>

<div id="uniquename2" style="display:none;">

The results from the study can be found in the `Friendly` data set in R after loading `library(car)`. 

Click the "Code" button to see the data.


```{r}
datatable(Friendly, options=list(lengthMenu = c(3,10,30)))
```


</div>
</div>

<br />


<!-- Begin writing your analysis below here. -->

<!-- Note that your goal is to use the Friendly data to show whether or not the Meshed or Before methods have any positive benefit on memory recall. -->

### What we asked

Does either approach prove to be beneficial? Do both approach's have the same results?
We want to know if we need to re-shape our curriculum to incorporate these strategies. If incorporating these strategies improves our students learning experience then we will include them right away!

We will perform three Wilcoxon Rank Sum Test's to answer these questions. 

<br />


#### Hypothesis

Our hypothesis for our first two tests. Here we are comparing the "Before Approach" and the "Meshed Approach" groups against the "SFR Approach" group.
$$
  H_0: \text{difference in medians} = 0
$$

$$
  H_a: \text{difference in medians} > 0
$$
<br />
Our hypothesis for our last test. Here we are comparing the "Before Approach" group and the "Meshed Approach" group.
$$
  H_0: \text{difference in medians} = 0
$$

$$
  H_a: \text{difference in medians} \neq 0
$$
For all three tests we'll use the following level of significance,
$$\alpha = 0.1$$

### Analysis

<br />
Did the results for the "Before Approach" group outperform the results of the "SFR" group? 
<br />See the results of the test below:

```{r}

# Is the Before approach better than the Standard Free Recall group?
wilcox.test(Friendly$correct[Friendly$condition == "Before"],Friendly$correct[Friendly$condition == "SFR"], mu=0, alternative="greater", conf.level=0.95) %>% 
  pander()
```

<br />
Did the results for the "Meshed Approach" group outperform the results of the "SFR" group?
<br />See the results of the test below:

```{r}

# Is the Meshed approach better than the Standard Free Recall group?
wilcox.test(Friendly$correct[Friendly$condition == "Meshed"],Friendly$correct[Friendly$condition == "SFR"], mu=0, alternative="greater", conf.level=0.95) %>%
  pander()
```

<br />
Are the results for the "Before Approach" group and the "Meshed Approach" group the same?
<br />See the results of the test below:
```{r}

# Are the benefits of the Meshed and Before groups the same? Let me phrase that better,
# Are the results for the Meshed and Before groups the same?
wilcox.test(Friendly$correct[Friendly$condition == "Meshed"],Friendly$correct[Friendly$condition == "Before"], mu=0, alternative="two.sided", conf.level=0.95) %>% 
  pander()
```


```{r}
#---------------------------------------------------------------------------------------------------------------#
#                           Tests with unexpected output:                                                       #
#---------------------------------------------------------------------------------------------------------------#

# This tests if the "Meshed" group has better results than the "Before" group. Ouput is True.

# wilcox.test(Friendly$correct[Friendly$condition == "Meshed"],Friendly$correct[Friendly$condition == "Before"], mu=0, alternative="greater", conf.level=0.95)


# This tests if the "Before" group has better results than the "Meshed" group. Output is True.

# wilcox.test(Friendly$correct[Friendly$condition == "Before"],Friendly$correct[Friendly$condition == "Meshed"], mu=0, alternative="greater", conf.level=0.95)

# How can both of those tests return True? 
#---------------------------------------------------------------------------------------------------------------#
```

### Conclusion

We asked the question, "Does either approach prove to be beneficial?", Its very important to note here that there are ties present in our data which adversely affects our results. However, our tests suggest that either approach proves to be beneficial for students



Do both approach's have the same results?


